```{r setup, echo=F}
library(knitr)
opts_chunk$set(echo=T, breaklines=T, fig.width=10, fig.path="figure-analysis/")
opts_knit$set(root.dir="../")
options(width=140)         
```
```{r echo=F, message=F, warning=F}
library(xtable)
library(ggplot2)
library(knitr)
```

# DMC2014 Data Analysis

## Read in the Data
```{r}
dt.train <- read.csv("task/orders_train.txt", sep=";", na.strings=c("?", "??", "NA", ""))
dt.test <- read.csv("task/orders_class.txt", sep=";", na.strings=c("?", "??", "NA", ""))

dt.dates <- c("dateOfBirth", "creationDate", "orderDate", "deliveryDate")
dt.train[dt.dates] <- lapply(dt.train[dt.dates], as.Date)
dt.test[dt.dates] <- lapply(dt.test[dt.dates], as.Date)
```

## General data Analysis

### Number of instances

```{r}
nrow(dt.train)
nrow(dt.test)
```

### Missing Values

```{r}
nas.train <- as.data.frame(apply(dt.train, 2, function(x) sum(is.na(x)) / length(x) * 100))
nas.test <- as.data.frame(apply(dt.test, 2, function(x) sum(is.na(x)) / length(x) * 100))
nas <- merge(nas.train, nas.test, by="row.names")
colnames(nas) <- c("Attribute", "Train (missing %)", "Test (missing %)")
kable(nas)
```

## Individual Attribute Analysis

### orderItemID

This item is unique and thus serves no purpose in the Machine Learning process:

```{r}
with(dt.train, length(unique(orderItemID)) / length(orderItemID))
with(dt.test, length(unique(orderItemID)) / length(orderItemID))
```

### orderDate

### deliveryDate

### itemID

## size

## color

## manufacturerID

## price

## customerID

## salutation

## dateOfBirth

We'll take a look at the histograms first:
```{r}
par(mfrow=c(1, 2))
barplot(table(dt.train$dateOfBirth), main="Histogram of dateOfBirth (Train)")
barplot(table(dt.test$dateOfBirth), main="Histogram of dateOfBirth (Test)")
```

Two points become obvious:
- Nobody lives that long.
- There is one DOB that appears unusually often in the two sets.

Which DOB appears that often?

```{r}
tail(sort(table(dt.train$dateOfBirth)))
tail(sort(table(dt.test$dateOfBirth)))
```

Set the dateOfBirth Attribute of these outliers to NA:

```{r}
outl.train <- with(dt.train, !is.na(dateOfBirth) & dateOfBirth %in% c(as.Date("1900-11-19"), as.Date("1949-11-19")))
outl.test <- with(dt.test, !is.na(dateOfBirth) & dateOfBirth %in% c(as.Date("1900-11-19"), as.Date("1949-11-19")))

dt.train[outl.train, ]$dateOfBirth <- NA
dt.test[outl.test, ]$dateOfBirth <- NA
```

Let's take a look again:

```{r}
par(mfrow=c(1, 2))
barplot(table(dt.train$dateOfBirth), main="Histogram of dateOfBirth (Train)")
barplot(table(dt.test$dateOfBirth), main="Histogram of dateOfBirth (Test)")
```

## state

## creationDate

## returnShipment