This document gets compiled automatically with each svn commit and is available at:
[http://dmc.0x0b.de](http://dmc.0x0b.de).

```{r setup, echo=F}
library(knitr)
opts_chunk$set(echo=F, breaklines=T)
opts_knit$set(fig.path="figure-notes/", root.dir="../")
```
```{r echo=F}
source("R/data.R")
source("R/dmc2010.R")
```
# DMC 2010 Notes

The task is to classify the target attribute ``target90`` (binary classification). If this
attribute is ``yes``, then no vouchers are given out and vice versa. So, the decision to
give out vouchers really is the inverse of ``target90``.

## Basic Numbers

### Maximum Points

The maximum achievable points:
```{r max}
points.max <- dmc.points(dt$target90, dt$target90)
points.max
```

### Baseline

The baseline can be constructed by giving out a voucher all the time:

```{r baseline}
points.baseline <- dmc.points(rep("no", nrow(dt)), dt$target90)
points.baseline
```

In terms of ratio:

```{r baseliner}
points.baseline / points.max
```


## Cost-Sensitivity

The problem at hand is a cost-sensitive classification problem, and
the cost matrix is given by

```{r cost}
dmc.cost
```

where columns correspond to the true classes and rows are the predicted class.
Diagonal elements are ignored, so it needs to be normalized.